---
title: Data 607 Project-3
author: 
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, cache=TRUE}
library(RMariaDB)
library(DBI)
library(bcputility)
library(tm)
library(dplyr)
library(stringr)
library(tidytext)
```

<H1> Data Read In </H1>
```{r, cache=TRUE}
#DB Set up 
con <- dbConnect(RMariaDB::MariaDB(), username="root", password="TestCase123.", dbname ="basic", host="localhost")
dbListTables(con)
```

```{r, cache=TRUE}
df <- dbReadTable(con, "Rectified_Total")
```

```{r}
sub_length = 100
new_df <- data.frame( doc_id = 1:sub_length, text =df$Job_Description[1:sub_length] )

```

<H1> Corpus Creation </H1>

```{r, cache=TRUE}
(ds <- DataframeSource(new_df))
v <- VCorpus(ds)
x <- Corpus(ds)

```
```{r, cache=TRUE}
writeCorpus(v, path = "./Vcorpus_sub")
writeCorpus(x, path = "./corpus_sub")
```

<H2> Load the corpus if needed </H2>

Start Here if you are pulling from git

```{r, cache=TRUE}
x <- Corpus(DirSource("./corpus_sub"), readerControl = list(language="lat"))
```


<H1> Let's Tidy it up and clean it </H1>

```{r, cache=TRUE}
inspect(x[1:2])
```
```{r, cache=TRUE}
x <- tm_map(x, content_transformer(tolower))
x <- tm_map(x, removeWords, stopwords("english"))
```

<H1> Let's see how our stripped instances are </H1>

```{r, cache=TRUE}
inspect(x[1:2])
```
<H1> Now Let's Create DTM </H1>

```{r, cache=TRUE}
dtm <- DocumentTermMatrix(x)
```

```{r, cache=TRUE}
inspect(dtm)
```
<H1> Fun Little Tidbits </H1>

Let's find the words asociated to other words that we are interested in! It could be all kinds of useful!

```{r, cache=TRUE}
findAssocs(dtm, "data", 0.40)
```

<H1> Let's Grab Technical Skills </H1>

<H2> The Master List </H2>

First things first, lets grab a master list of coding languages. 

```{r, cache=TRUE}
url.data <- "https://raw.githubusercontent.com/jamhall/programming-languages-csv/master/languages.csv"
raw <- read.csv(url(url.data), header = TRUE,)
programming_list <- tolower(raw$name)
```

<H2> Technical Skill data </H2>

Now lets see which jobs match which roles

```{r, cache=TRUE}
programming_list_dtm <- DocumentTermMatrix(x, list(dictionary = c(programming_list)))
```

```{r, cache=TRUE}
programming_list_df <- as.data.frame(as.matrix(programming_list_dtm), stringsAsFactors=False)
```



<H2> Let Get Better Metrics Counts </H2>

Let's first test a routine, we're going to try and count the number of rows that exist for python.


```{r, cache=TRUE}
sum(programming_list_df$python != 0, na.rm=TRUE)
```
<H2> Now Let's Apply it elsewhere </H2>

Now let make a new dataframe. First things first, let's use this as a way to get skills and counts. 

```{r, cache=TRUE}
index_df <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(index_df) <-c("Skill", "Count")

```

From there, let's populate it

```{r, cache=TRUE}
list_holder <- list()
for(i in 1:ncol(programming_list_df)) {       # for-loop over columns
  index_df[i,] <- c(colnames(programming_list_df)[i], (sum(programming_list_df[, i] != 0, na.rm=TRUE)))
}
index_df
```

```{r, cache=TRUE}
list_holder <- list()
for(i in 1:ncol(programming_list_df)) {       # for-loop over columns
  index_df[i,] <- c(colnames(programming_list_df)[i], (sum(programming_list_df[, i] != 0, na.rm=TRUE)))
}
index_df
```
Now let's remove everything that is useless. 

```{r, cache=TRUE}

finalizedProgramList <- index_df[index_df$Count !=0,]
```

And finally, lets view our data. Its a lot smaller!


```{r, cache=TRUE}
finalizedProgramList
```

<H1> The Non-techncial skills </H1>

First things first, lets grab a master list of coding languages. 

```{r, cache=TRUE}
url.data <- "https://raw.githubusercontent.com/Amantux/Project_3/main/Data_Skills.csv"
raw <- read.csv(url(url.data), header= TRUE,)
head(raw)
soft_skills <- tolower(raw$ï..Skills) #Very confused at the column name, but I'm just rolling with it
```

<H2> Non-Technical Skill data </H2>

Now lets see which jobs match which roles

```{r, cache=TRUE}
soft_skills_dtm <- DocumentTermMatrix(x, list(dictionary = c(soft_skills)))
```

```{r, cache=TRUE}
soft_skills
```


```{r, cache=TRUE}
inspect(soft_skills_dtm)
```



```{r, cache=TRUE}
soft_skills_dtm <- as.data.frame(as.matrix(soft_skills_dtm), stringsAsFactors=False)
```



<H2> Let Get Better Metrics Counts </H2>

Let's first test a routine, we're going to try and count the number of rows that exist for python.


```{r, cache=TRUE}
sum(soft_skills_dtm$analytics != 0, na.rm=TRUE)
```
<H2> Now Let's Apply it elsewhere </H2>

Now let make a new dataframe. First things first, let's use this as a way to get skills and counts. 

```{r, cache=TRUE}
index_df_soft <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(index_df_soft) <-c("Skill", "Count")

```

From there, let's populate it

```{r, cache=TRUE}
list_holder <- list()
for(i in 1:ncol(soft_skills_dtm)) {       # for-loop over columns
  index_df_soft[i,] <- c(colnames(soft_skills_dtm)[i], (sum(soft_skills_dtm[, i] != 0, na.rm=TRUE)))
}
index_df_soft
```

```{r, cache=TRUE}
list_holder <- list()
for(i in 1:ncol(soft_skills_dtm)) {       # for-loop over columns
  index_df_soft[i,] <- c(colnames(soft_skills_dtm)[i], (sum(soft_skills_dtm[, i] != 0, na.rm=TRUE)))
}
index_df_soft
```
Now let's remove everything that is useless. 

```{r, cache=TRUE}

finalizedSoftList <- index_df_soft[index_df$Count !=0,]
```

And finally, lets view our data. Its a lot smaller!


```{r, cache=TRUE}
finalizedSoftList
```
